---
title: "Serova_For_Report"
author: "Serova"
date: "2025-06-18"
output: html_document
---

# loading all required libraries
```{r}

library(tidyr)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(randomForest)
library(tidyr)
library(lattice)
library(xgboost)  
library(caret)
library(Matrix)
library(dplyr)
library(glmnet)
library(e1071)
library(pROC)
library(caret)

```



```{r}
data <- read.csv("LDL_students_scores+meta+wei+hei.csv", header = TRUE)

# Preprocessing
data <- data %>%
  filter(unit == "mmol/L") %>%
  mutate(
    BMI = ifelse(height == 0, 0, weight / (height / 100)^2)
  ) %>%
  filter(between(height, 100, 200),
         between(age.x, 18, 70),
         between(weight, 40, 160),
         between(value, 0.5, 7),
         between(BMI, 16, 40)) %>%
  filter(!is.na(smoking_status) & !is.na(alcoholVolume) & !is.na(alcohol_consumption))

data$smoking_status <- as.character(data$smoking_status)
data$smoking_status[data$smoking_status == ""] <- NA
data <- data[!is.na(data$smoking_status), ]


# Convert selected columns to factors
cols_to_factor <- c("alcohol_consumption", "alcoholVolume", "smoking_status", "gender.x")
data[cols_to_factor] <- lapply(data[cols_to_factor], as.factor)

# Remove duplicates
data <- data[!duplicated(data[, -1]), ]

# List of features
features <- c("score", "gender.x", "age.x", "alcohol_consumption", 
              "smoking_status", "alcoholVolume", "weight", "height", "BMI")

# Generate feature combinations
combinations <- list()
for (i in 2:length(features)) {
  combinations <- c(combinations, lapply(combn(features, i, simplify = FALSE), as.vector))
}

# Model evaluation function
evaluate_models <- function(data, combination) {
  set.seed(42)
  train_indices <- sample(1:nrow(data), size = 0.8 * nrow(data))
  train_data <- data[train_indices, ]
  test_data <- data[-train_indices, ]
  
  train_data <- na.omit(train_data)
  test_data <- na.omit(test_data)

  # Drop unused factor levels in training set
  train_data <- droplevels(train_data)

  # Synchronize factor levels in test set with train set
  for (col in combination) {
    if (is.factor(train_data[[col]])) {
      test_data[[col]] <- factor(test_data[[col]], levels = levels(train_data[[col]]))
    }
  }

  # Prepare model matrices
  train_matrix <- model.matrix(~ . - 1, data = train_data[, combination])
  test_matrix <- model.matrix(~ . - 1, data = test_data[, combination])
  preProc <- preProcess(train_matrix, method = c("center", "scale"))
  train_scaled <- predict(preProc, train_matrix)
  test_scaled <- predict(preProc, test_matrix)
  
  # Linear regression
  model_lm <- glm(value ~ ., data = train_data[, c("value", combination)])
  pred_lm <- predict(model_lm, newdata = test_data)
  
  # Random Forest
  model_rf <- randomForest(value ~ ., data = train_data[, c("value", combination)],
                           ntree = 300, mtry = max(floor(length(combination)/3), 1),
                           nodesize = 5, importance = TRUE)
  pred_rf <- predict(model_rf, newdata = test_data)
  
  if (nrow(train_scaled) != length(train_data$value) || nrow(test_scaled) != length(test_data$value)) {
  warning("Dimensions do not match, skipping combination: ", paste(combination, collapse = ", "))
  return(NULL)
}
  
  
  # XGBoost
  dtrain <- xgb.DMatrix(data = train_scaled, label = train_data$value)
  dtest <- xgb.DMatrix(data = test_scaled, label = test_data$value)
  xgb_model <- xgb.train(params = list(objective = "reg:squarederror", eta = 0.01,
                                       max_depth = 3, subsample = 0.7, colsample_bytree = 0.7,
                                       lambda = 2, alpha = 1, verbosity = 0),
                         data = dtrain, nrounds = 300,
                         early_stopping_rounds = 20,
                         watchlist = list(train = dtrain, test = dtest), verbose = 0)
  pred_xgb <- predict(xgb_model, dtest)
  
  # SVR
  model_svr <- svm(value ~ ., data = train_data[, c("value", combination)])
  pred_svr <- predict(model_svr, newdata = test_data)
  
  # Elastic Net
  model_en <- cv.glmnet(train_scaled, train_data$value, alpha = 0.5)
  pred_en <- predict(model_en, s = "lambda.min", newx = test_scaled)
  
  # Evaluation metrics
  get_metrics <- function(actual, pred) {
    c(MAE = mean(abs(actual - pred), na.rm = TRUE),
      MSE = mean((actual - pred)^2, na.rm = TRUE),
      R2 = cor(actual, pred, use = "complete.obs")^2)
  }
  
  data.frame(
    Combination = paste(combination, collapse = ", "),
    Model = c("Linear Regression", "Random Forest", "XGBoost", "SVR", "Elastic Net"),
    rbind(
      get_metrics(test_data$value, pred_lm),
      get_metrics(test_data$value, pred_rf),
      get_metrics(test_data$value, pred_xgb),
      get_metrics(test_data$value, pred_svr),
      get_metrics(test_data$value, pred_en)
    )
  )
}

# Get evaluation results
results <- do.call(rbind, lapply(combinations, function(comb) {
  cat("Processing:", paste(comb, collapse = ", "), "\n")
  evaluate_models(data, comb)
}))

```


# Histogram of age
```{r}
hist(data_no_dup_filt_age$age.x,   
     main = "Age",     
     xlab = "Years",               
     ylab = "Frequency",                 
     breaks = 20)  
grid()
```

# Histogram of height
```{r}
hist(data_no_dup_filt_age$height,   
     main = "Height",     
     xlab = "cm",               
     ylab = "Frequency",                 
     breaks = 30)  
grid()
```


# Histogram of weight
```{r}
hist(data_no_dup_filt_age_weight$weight,   
     main = "Weight",     
     xlab = "kg",               
     ylab = "Frequency",                 
     breaks = 30)  
grid()
```

# Histogram of BMI
```{r}
hist(data_no_dup_filt_age_weight_bmi$BMI,   
     main = "BMI",     
     xlab = "",               
     ylab = "Frequency",                 
     breaks = 20)  
grid()
```

# Histogram of LDL value
```{r}
hist(data_no_dup_filt_value$value,   
     main = "LDL value",     
     xlab = "value, mmol/L",               
     ylab = "Frequency",                 
     breaks = 30)  
grid()
```

# Histogram of PRS
```{r}
hist(data_no_dup_filt_value$score,   
     main = "Polygenic Risk Score",     
     xlab = "PRS",               
     ylab = "Frequency",                 
     breaks = 30)  
grid()
```



#Cross-validation of top-5 models

```{r}
top_models <- results %>% arrange(desc(R2)) %>% head(5)


cross_validate_model <- function(data, combination, model_type, n_folds = 5) {
  set.seed(42)
  
  data <- data[!is.na(data$value), ]

  
  folds <- createFolds(data$value, k = n_folds, list = TRUE)
  metrics_list <- list()
  
  for (i in seq_along(folds)) {
    train_indices <- unlist(folds[-i])
    test_indices <- folds[[i]]
    train_data <- data[train_indices, ]
    test_data <- data[test_indices, ]
    
    train_matrix <- model.matrix(~ . - 1, data = train_data[, combination])
    test_matrix <- model.matrix(~ . - 1, data = test_data[, combination])
    preProc <- preProcess(train_matrix, method = c("center", "scale"))
    train_scaled <- predict(preProc, train_matrix)
    test_scaled <- predict(preProc, test_matrix)
    
    if (model_type == "Linear Regression") {
      model <- glm(value ~ ., data = train_data[, c("value", combination)])
      pred <- predict(model, newdata = test_data)
    } else if (model_type == "Random Forest") {
      model <- randomForest(value ~ ., data = train_data[, c("value", combination)],
                            ntree = 300, mtry = max(floor(length(combination)/3), 1),
                            nodesize = 5)
      pred <- predict(model, newdata = test_data)
    } else if (model_type == "XGBoost") {
      dtrain <- xgb.DMatrix(data = train_scaled, label = train_data$value)
      dtest <- xgb.DMatrix(data = test_scaled, label = test_data$value)
      model <- xgb.train(params = list(objective = "reg:squarederror", eta = 0.01,
                                       max_depth = 3, subsample = 0.7, colsample_bytree = 0.7,
                                       lambda = 2, alpha = 1, verbosity = 0),
                         data = dtrain, nrounds = 300, verbose = 0)
      pred <- predict(model, dtest)
    } else if (model_type == "SVR") {
      model <- svm(value ~ ., data = train_data[, c("value", combination)])
      pred <- predict(model, newdata = test_data)
    } else if (model_type == "Elastic Net") {
      model <- cv.glmnet(train_scaled, train_data$value, alpha = 0.5)
      pred <- predict(model, s = "lambda.min", newx = test_scaled)
    }
    
    metrics <- c(MAE = mean(abs(test_data$value - pred), na.rm = TRUE),
                 MSE = mean((test_data$value - pred)^2, na.rm = TRUE),
                 R2 = cor(test_data$value, pred, use = "complete.obs")^2)
    metrics_list[[i]] <- metrics
  }
  
  colMeans(do.call(rbind, metrics_list))
}


cv_results <- do.call(rbind, lapply(1:nrow(top_models), function(i) {
  row <- top_models[i, ]
  combination <- unlist(strsplit(row$Combination, ", "))
  model_type <- row$Model
  metrics <- cross_validate_model(data, combination, model_type)
  c(Combination = row$Combination, Model = model_type, metrics)
}))

print(cv_results)
```


#Building a Random Forest model with the best combination of predictors.
```{r}

set.seed(42)
train_indices_noalco <- sample(1:nrow(df_clean_noalco), size = 0.8 * nrow(df_clean_noalco))  
train_data_rf_noalco <- df_clean_noalco[train_indices_noalco, ]  
test_data_rf_noalco <- df_clean_noalco[-train_indices_noalco, ]

train_data_rf_noalco <- na.omit(train_data_rf_noalco)
test_data_rf_noalco <- na.omit(test_data_rf_noalco)

rf_model_noalco <- randomForest(value ~ score + age.x + alcohol_consumption + smoking_status + alcoholVolume + BMI, 
                         data = train_data_rf_noalco, 
                         ntree = 300,  
                         mtry = 2,    
                         importance = TRUE) 

rf_predictions_no_alco <- predict(rf_model_noalco, newdata = test_data_rf_noalco)
rf_r_squared_noalco <- cor(test_data_rf_noalco$value, rf_predictions_no_alco)^2
cat("R² (Random Forest):", round(rf_r_squared_noalco, 3))
```


# MSE и MAE:

```{r}
mse_rf_noalco <- mean((test_data_rf_noalco$value - rf_predictions_no_alco)^2)
cat("MSE:", mse_rf_noalco, "\n")    

mae_rf_noalco <- mean(abs(test_data_rf_noalco$value - rf_predictions_no_alco))
cat("MAE:", mae_rf_noalco, "\n")

mape_rf_noalco <- mean(abs((test_data_rf_noalco$value - rf_predictions_no_alco) / test_data_rf_noalco$value)) * 100
cat("MAPE:", mape_rf_noalco, "\n")
```



#Actual and predicted value
```{r}
# Set the plot to be square
par(pty = "s")

# Create a plot with equal limits on both axes
plot(test_data_rf_noalco$value, rf_predictions_no_alco,
     xlim = range(c(test_data_rf_noalco$value, rf_predictions_no_alco)),
     ylim = range(c(test_data_rf_noalco$value, rf_predictions_no_alco)),
     xlab = "Actual value", 
     ylab = "Predicted value", 
     main = "Actual vs Predicted values",
     pch = 19, col = "skyblue")  

# Add a diagonal line passing through 0, with slope 1
abline(a = 0, b = 1, col = "red")  

# Add grid lines
grid()
```

#Residual Plot

```{r}
residuals <- test_data_rf_noalco$value - rf_predictions_no_alco

# Plot settings
par(mfrow = c(1,1), pty = "m", mar = c(4,4,3,2))

# Create an improved plot
plot(rf_predictions_no_alco, residuals,
     xlab = "Predicted Values", 
     ylab = "Residuals (Actual - Predicted)",
     main = "Residuals Analysis of Regression Model",
     pch = 19, col = adjustcolor("steelblue", alpha.f = 0.6),
     cex = 1.2, las = 1,
     ylim = c(-max(abs(residuals))*1.1, max(abs(residuals))*1.1))

# Add elements
abline(h = 0, col = "red3", lwd = 2)  # Horizontal zero line
grid(col = "lightgray", lty = "dotted")  # Grid for better readability

# Add a smoothing line
lines(lowess(rf_predictions_no_alco, residuals), col = "darkgreen", lwd = 2)

# Add a legend
legend("topright", 
       legend = c("Residuals", "Zero Line", "Residuals Trend"),
       col = c("steelblue", "red3", "darkgreen"),
       pch = c(19, NA, NA), lty = c(NA, 1, 1), lwd = 2)
```


#Obtaining the importance of each predictor
```{r}
print(importance(rf_model_noalco))
```


# Feature importance plot:

```{r}
varImpPlot(rf_model_noalco, 
           main = "Importance of predictors", 
           type = 1)  # type = 1 for Mean Decrease Accuracy
```


# Plot of LDL dependence on PRS by age and gender:
```{r}
# Define age groups (in 5-year intervals)
age_breaks <- seq(20, 80, by = 5)
age_labels <- paste(head(age_breaks, -1), tail(age_breaks, -1), sep = "-")

# Add an age group column
df_clean_noalco <- df_clean_noalco %>%
  mutate(age_group = cut(
    age.x,
    breaks = age_breaks,
    labels = age_labels,
    right = FALSE,
    include.lowest = TRUE
  )) %>%
  filter(!is.na(age_group))

# Create a boxplot highlighting the normal zone and differentiating by gender
ggplot(df_clean_noalco, aes(x = age_group, y = value, fill = gender.x)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = 1.97, ymax = 4.65,
           fill = "lightgreen", alpha = 0.3) +
  geom_boxplot(outlier.alpha = 0.2) +
  labs(
    title = "The influence of age on LDL levels by age groups",
    x = "Age group",
    y = "LDL",
    fill = "Gender"
  ) +
  scale_fill_manual(values = c("lightpink", "lightblue")) +
  theme_minimal(base_size = 20) +
  theme(
    axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    plot.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 24),
    legend.title = element_text(size = 26)
  )
```


# Plot of value dependence on score:

```{r}
ggplot(df_clean_noalco, aes(x = score, y = value)) +  
  geom_point(shape = 21,                     # Point shape
             color = "blue",                 # Outline color
             fill = "skyblue",               # Fill color
             size = 3) +                     # Point size
  geom_smooth(method = "lm", color = "red") +   # Linear model fit line
  labs(title = "LDL Dependence on PRS",        # Plot title
       x = "PRS",                              # X-axis label
       y = "LDL") +                            # Y-axis label
  theme(
    # Increase axis text size
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    
    # Increase axis labels size
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    
    # Increase plot title size
    plot.title = element_text(size = 20, face = "bold")
  )
```


#Comparison of actual and predicted LDL values

```{r}
# Add predicted values to the dataset
test_data_rf_noalco$predicted_value <- rf_predictions_no_alco

# Define age intervals (5-year ranges)
age_breaks <- c(20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70)

# Generate interval labels like "20-25", "25-30", etc.
age_labels <- paste(head(age_breaks, -1), tail(age_breaks, -1), sep = "-")

# Assign age groups based on intervals
test_data_rf_noalco <- test_data_rf_noalco %>%
  mutate(age_group = cut(
    age.x,
    breaks = age_breaks,
    labels = age_labels,
    right = FALSE,
    include.lowest = TRUE
  ))

library(tidyr)

# Convert data to long format for plotting
plot_data <- test_data_rf_noalco %>%
  pivot_longer(
    cols = c(value, predicted_value),
    names_to = "value_type",
    values_to = "LDL_level"
  )

# Plot comparison with boxplots and facets by gender
ggplot(plot_data, aes(x = age_group, y = LDL_level, fill = value_type)) +
  annotate("rect", 
           xmin = -Inf, xmax = Inf, 
           ymin = 1.91, ymax = 4.65,
           fill = "lightgreen", alpha = 0.3) +
  geom_boxplot(alpha = 0.7, color = "grey30") +
  labs(
    title = "Comparison of Actual and Predicted LDL Values",
    x = "Age Group",
    y = "LDL Level (mmol/L)"
  ) +
  scale_fill_manual(
    values = c("value" = "#1f78b4", "predicted_value" = "#33a02c"),
    labels = c("Actual Value", "Predicted Value"),
    name = "Measurement Type"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    plot.title = element_text(size = 20, face = "bold"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    strip.text = element_text(size = 18, face = "bold")  # For facet labels
  ) +
  facet_wrap(~ gender.x, ncol = 2)
```

# Look at the percentage of people outside the normal range in PRS bins:
```{r}
# LDL norm boundaries
lower_norm <- 1.97
upper_norm <- 4.65

# 1. Create 10% PRS bins (10 groups)
df_clean_noalco <- df_clean_noalco %>%
  mutate(
    PRS_group = cut(
      score,
      breaks = quantile(score, probs = seq(0, 1, 0.1), na.rm = TRUE),
      labels = paste0(seq(0, 90, 10), "-", seq(10, 100, 10), "%"),
      include.lowest = TRUE
    )
  ) %>%
  filter(!is.na(PRS_group))

# 2. Calculate LDL deviation percentages
result <- df_clean_noalco %>%
  group_by(PRS_group) %>%
  summarize(
    percent_below = mean(value < lower_norm, na.rm = TRUE) * 100,
    percent_above = mean(value > upper_norm, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# 3. Prepare data for plotting
result_long <- result %>%
  pivot_longer(
    cols = c(percent_below, percent_above),
    names_to = "deviation_type",
    values_to = "percent"
  ) %>%
  mutate(
    deviation_type = recode(deviation_type,
                          "percent_below" = "Below normal",
                          "percent_above" = "Above normal")
  )

# 4. Create the plot
ggplot(result_long, aes(x = PRS_group, y = percent, fill = deviation_type)) +
  geom_bar(stat = "identity", alpha = 0.8, width = 0.7) +
  labs(
    title = "Percentage of People with LDL Deviations by PRS Deciles",
    x = "PRS Groups (Deciles)",
    y = "Percentage of People (%)",
    fill = "Deviation Category"
  ) +
  scale_fill_manual(
    values = c("Below normal" = "#FF6B6B", "Above normal" = "#6B8EFF"),
    labels = c("Below normal", "Above normal")
  ) +
  theme_minimal(base_size = 16) +  # базовый размер текста
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    legend.position = "top",
    panel.grid.major.x = element_blank()
  )
```


#The same, but for better visualization, we will split the plots for values above the normal range and below the normal range.

```{r}
ggplot(result_long, aes(x = PRS_group, y = percent, fill = deviation_type)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_text(
    aes(label = sprintf("%.1f%%", percent)),
    vjust = -0.5,
    size = 5,  # Increase text size above columns
    color = "black"
  ) +
  facet_wrap(~deviation_type, scales = "free_y") +
  scale_fill_manual(values = c("Below normal" = "#6B8EFF", "Above normal" = "#FF6B6B")) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.15))  # 15% padding on top
  ) +
  labs(
    title = "LDL Deviations Out of Norm by PRS Deciles",
    x = "PRS Groups (Deciles)",
    y = "Percentage of People (%)"
  ) +
  theme_minimal(base_size = 16) +  # Global font size increase
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 16, face = "bold"),  # Facet titles
    legend.position = "none",
    panel.grid.major.x = element_blank()
  )
```

# Now, split the lower and upper PRS groups into percentiles and then estimate their Odds Ratios (OR) within each group.
```{r}
# 1. Create custom PRS bins (0-5, 5-10, 10-20, ..., 90-95, 95-100)
prs_breaks <- quantile(df_clean_noalco$score, probs = c(0, 0.05, 0.10, seq(0.2, 0.9, 0.1), 0.95, 1.0), na.rm = TRUE)
prs_labels <- c("0-5%", "5-10%", "10-20%", "20-30%", "30-40%", "40-50%", 
                "50-60%", "60-70%", "70-80%", "80-90%", "90-95%", "95-100%")

df_clean_noalco <- df_clean_noalco %>%
  mutate(
    PRS_group = cut(score, breaks = prs_breaks, labels = prs_labels, include.lowest = TRUE)
  ) %>%
  filter(!is.na(PRS_group))

# 2. Calculate LDL deviation percentages
result <- df_clean_noalco %>%
  group_by(PRS_group) %>%
  summarize(
    percent_below = mean(value < lower_norm, na.rm = TRUE) * 100,
    percent_above = mean(value > upper_norm, na.rm = TRUE) * 100,
    .groups = "drop"
  )

# 3. Prepare data for plotting
result_long <- result %>%
  pivot_longer(
    cols = c(percent_below, percent_above),
    names_to = "deviation_type",
    values_to = "percent"
  ) %>%
  mutate(
    deviation_type = recode(deviation_type,
                            "percent_below" = "Below normal",
                            "percent_above" = "Above normal")
  )

# 4. Create the plot

ggplot(result_long, aes(x = PRS_group, y = percent, fill = deviation_type)) +
  geom_col(alpha = 0.8, width = 0.7) +
  scale_fill_manual(
    values = c("Below normal" = "#FF6B6B", "Above normal" = "#6B8EFF")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Percentage of People with LDL Deviations by PRS Percentiles",
    x = "PRS Groups (Percentiles)",
    y = "Percentage of People (%)",
    fill = "Deviation Category"
  ) +
  theme_minimal(base_size = 20) +  # увеличен базовый размер
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 18),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 18),
    legend.position = "top",
    panel.grid.major.x = element_blank()
  )

```

# Plot of OR (Odds Ratio) for each percentile group among those below the normal range

```{r}
# 1. Set the normal range boundary
lower_norm <- 1.97

# 2. Create the target variable: below normal = 1, normal and above = 0
df_clean_noalco <- df_clean_noalco %>%
  mutate(ldl_below = ifelse(value < lower_norm, 1, 0))

# 3. Create PRS deciles
labels_deciles <- paste0(seq(0, 90, 10), "-", seq(10, 100, 10), "%")
df_clean_noalco <- df_clean_noalco %>%
  mutate(PRS_decile = cut(
    score,
    breaks = quantile(score, probs = seq(0, 1, 0.1), na.rm = TRUE),
    labels = labels_deciles,
    include.lowest = TRUE
  ))

# 4. Check for the presence of the required reference group
decile_levels <- levels(df_clean_noalco$PRS_decile)
ref_group <- if ("0-5%" %in% decile_levels) "0-5%" else "0-10%"

# 5. Set the reference group
df_clean_noalco$PRS_decile <- relevel(factor(df_clean_noalco$PRS_decile), ref = ref_group)

# 6. Logistic regression model
model <- glm(ldl_below ~ PRS_decile, data = df_clean_noalco, family = binomial)

# 7. Get OR and confidence intervals
or_data <- as.data.frame(exp(cbind(OR = coef(model), confint(model))))
or_data$Decile <- rownames(or_data)

# 8. Keep only the decile coefficients
or_data <- or_data[grep("PRS_decile", or_data$Decile), ]
or_data$Decile <- gsub("PRS_decile", "", or_data$Decile)

# 9. Define the order of levels
or_data$Decile <- factor(or_data$Decile, levels = labels_deciles)

# 10. Add a row for the reference group
ref_row <- data.frame(
  OR = 1,
  `X2.5.` = NA_real_,
  `X97.5.` = NA_real_,
  Decile = factor(ref_group, levels = labels_deciles)
)

# Rename columns
names(or_data)[names(or_data) == "2.5 %"] <- "X2.5."
names(or_data)[names(or_data) == "97.5 %"] <- "X97.5."

# 11. Combine the data
or_data <- rbind(ref_row, or_data)

# 12. Final plot
library(ggplot2)
plot <- ggplot(or_data, aes(x = Decile, y = OR)) +
  geom_point(size = 5, color = "#1f78b4", shape = 16) +
  geom_errorbar(aes(ymin = X2.5., ymax = X97.5.),
                width = 0.4, color = "#1f78b4", linewidth = 1.4) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 1) +
  labs(
    title = "Odds Ratio of LDL Deficiency by PRS Decile",
    subtitle = paste("Reference group:", ref_group),
    x = "PRS deciles",
    y = "Odds Ratio (Below norm)"
  ) +
  theme_minimal(base_size = 17) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 13),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 20),
    plot.subtitle = element_text(hjust = 0.5, size = 15),
    axis.title = element_text(size = 15)
  )

# 13. Save the plot
ggsave("LDL_OR_plot_ref_lowest.png", plot = plot, width = 16, height = 12, units = "in", dpi = 300)
```


# Plot of OR (Odds Ratio) for each percentile group among those above the normal range
```{r}
# 1. Set the upper limit of the normal range
upper_norm <- 4.65

# 2. Target variable: above normal = 1, normal and below = 0
df_clean_noalco <- df_clean_noalco %>%
  mutate(ldl_above = ifelse(value > upper_norm, 1, 0))

# 3. Create PRS deciles
labels_deciles <- paste0(seq(0, 90, 10), "-", seq(10, 100, 10), "%")
df_clean_noalco <- df_clean_noalco %>%
  mutate(PRS_decile = cut(
    score,
    breaks = quantile(score, probs = seq(0, 1, 0.1), na.rm = TRUE),
    labels = labels_deciles,
    include.lowest = TRUE
  ))

# 4. Set reference group
ref_group <- "0-10%"
df_clean_noalco$PRS_decile <- relevel(factor(df_clean_noalco$PRS_decile), ref = ref_group)

# 5. Logistic regression model
model_above <- glm(ldl_above ~ PRS_decile, data = df_clean_noalco, family = binomial)

# 6. OR and confidence intervals
or_above <- as.data.frame(exp(cbind(OR = coef(model_above), confint(model_above))))
or_above$Decile <- rownames(or_above)
or_above <- or_above[grep("PRS_decile", or_above$Decile), ]
or_above$Decile <- gsub("PRS_decile", "", or_above$Decile)

# 7. Set order of levels
or_above$Decile <- factor(or_above$Decile, levels = labels_deciles)

# 8. Rename columns
colnames(or_above)[colnames(or_above) == "2.5 %"] <- "CI_low"
colnames(or_above)[colnames(or_above) == "97.5 %"] <- "CI_high"

# 9. Add a row for the reference group with OR=1 and CI=1
ref_above <- data.frame(
  OR = 1,
  CI_low = 1,
  CI_high = 1,
  Decile = factor(ref_group, levels = labels_deciles)
)

# 10. Combine the reference with the OR data
or_above <- rbind(ref_above, or_above)

# 11. Plot
ggplot(or_above, aes(x = Decile, y = OR)) +
  geom_point(color = "#E31A1C", size = 5, shape = 16) +
  geom_errorbar(aes(ymin = CI_low, ymax = CI_high),
                width = 0.4, color = "#E31A1C", linewidth = 1.4) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 1) +
  labs(
    title = "Odds Ratio of LDL Excess by PRS Decile",
    subtitle = paste0("Reference group: ", ref_group),
    x = "PRS deciles",
    y = "Odds Ratio (Above norm)"
  ) +
  theme_minimal(base_size = 17) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    axis.text.y = element_text(size = 13),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 20),
    plot.subtitle = element_text(hjust = 0.5, size = 15),
    axis.title = element_text(size = 15)
  )

# 12. Save the plot
ggsave("LDL_OR_plot_ref_0_10_above.png", width = 16, height = 12, units = "in", dpi = 300)
```

